---
title: "Visualization"
author: "Lauren Holley"
date: 2026-09-17
output: github_document
---
```{r message=FALSE}
library(tidyverse)
library(ggridges)

library(p8105.datasets)
data("weather_df")
```

Now we have everything we need to start!

```{r}
weather_df
```

Let's make a scatterplot!

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax))
```

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```

Dataframe first typically...

```{r}
ggp_temp_scatterplot = 
  weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

ggp_temp_scatterplot
```

Let's make fancy woohoo

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.25) +
  geom_smooth(se = FALSE)
```

`alpha` adds some transparency into the points
`color` colors by name of area
`geom_smooth` adds smooth line through each set of interest
`se + FALSE` removes error bars

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = 0.25) +
  geom_smooth(se = FALSE)
```

Defining the aesthetic in a specific area so you can layer on the aesthetics.

The aesthetics are up to you!!!

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_smooth(se = FALSE)
```

Show faceting

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  facet_grid(. ~ name)
```

Second value (in this case name) will define columns in dataset with `facet_grid`.
The first value will separate the rows for the graph. If we don't want anything in there use `.` 

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  facet_grid(cols = vars(name))
```

Diff notation of the same thing tbh.

Let's look at something else neow

```{r}
weather_df |>
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = 0.5) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

Make a plot of central park tmax v tmin only, and convert temperatures to farenheit.

```{r}
weather_df |>
  filter(name == "CentralPark_NY") |>
  mutate(
    tmax = tmax * (9 / 5) + 32,
    tmin = tmin * (9 / 5) + 32
  ) |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

what is a hex plot

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex()
```

## Univariate Plots

```{r}
weather_df |>
  ggplot(aes(x = tmax)) +
  geom_histogram()
```

```{r}
weather_df |>
  ggplot(aes(x = tmax, fill = name)) +
  geom_histogram() +
  facet_grid(. ~ name)
```

`position = "dodge"` in `geom_histogram` means that the bars won't overlap
`fill` instead of `color` to fill bars.

Density plots are great

```{r}
weather_df |>
  ggplot(aes(x = tmax, fill = name)) +
  geom_density(alpha = 0.3)
```

Boxplots

```{r}
weather_df |>
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()
```

Violin plot hehe

```{r}
weather_df |>
  ggplot(aes(x = name, y = tmax)) +
  geom_violin()
```

Ridge plot

```{r}
weather_df |>
  ggplot(aes(x = tmax, y = name)) +
  geom_density_ridges()
```

## Save some of my plots

```{r}
ggp_weather =
  weather_df |>
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = 0.5) +
  facet_grid(. ~ name)

ggp_weather

ggsave("ggp_weather.pdf", ggp_weather)
```

```{r fig.width = 12, fig.height = 12}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

