---
title: "Visualization Part 2"
author: "Lauren Holley"
date: 2026-10-06
output: github_document
---
```{r message=FALSE}
library(tidyverse)
library(patchwork)
library(p8105.datasets)
data("weather_df")
```

Starting with scatterplot:

```{r}
weather_df |>
  ggplot(aes(x = tmax, y = tmin, color = name)) +
  geom_point() +
  labs(
    title = "Temperature (Max vs Min)",
    x = "Max Temperature (C)",
    y = "Min Temperature (C)",
    color = "Location",
    caption = "Data from NOAA for three weather stations."
  )
```

Let's try some other scales

```{r}
weather_df |>
  ggplot(aes(x = tmax, y = tmin, color = name)) +
  geom_point() +
  labs(
    title = "Temperature (Max vs Min)",
    x = "Max Temperature (C)",
    y = "Min Temperature (C)",
    color = "Location",
    caption = "Data from NOAA for three weather stations."
  ) + 
  scale_x_continuous(
    breaks = c(-10, 0, 15),
    labels = c("-10 C", "0", "Fifteen")
  ) +
  scale_y_continuous(
    trans = "sqrt",
    position = "right"
  )
```

`breaks` creates specific markings for the x axis labels/numbers

Let's look at color

```{r}
weather_df |>
  ggplot(aes(x = tmax, y = tmin, color = name)) +
  geom_point() +
  scale_color_hue(h = c(150, 600))
```

```{r}
weather_df |>
  ggplot(aes(x = tmax, y = tmin, color = name)) +
  geom_point() +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE
  )
```

## Themes

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE
  ) +
  theme(legend.position = "bottom")
```

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

`bw` black + white bars
`minimal` no outer outline
`classic` no lines

all just through a `ggthemes` package — can look up other options online

Update the tmax vs date plot.

```{r}
weather_df |>
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(alpha = 0.65) +
  geom_smooth(se = FALSE) +
  labs(
    title = "Seasonal Trends in Max Temperature",
    x = "Date",
    y = "Max Temperature (C)",
    caption = "Max daily temp in three weather stations in 2021 and 2022",
    color = "Location"
  ) +
  viridis::scale_color_viridis(
    discrete = TRUE
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

## Two more weird but useful plot things

```{r}
central_park_df =
  weather_df |>
  filter(name == "CentralPark_NY")

molokai_df =
  weather_df |>
  filter(name == "Molokai_HI")

ggplot(molokai_df, aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_line(data = central_park_df)
```

Multiple panels with different plot types.

```{r}
ggp_tmax_tmin = 
  weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() +
  theme(legend.position = "none")

ggp_prcp_density = 
  weather_df |>
  filter(prcp > 0) |>
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = 0.5) +
  theme(legend.position = "none")

ggp_seasonal = 
  weather_df |>
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  theme(legend.position = "bottom")

(ggp_tmax_tmin + ggp_prcp_density) / ggp_seasonal
```

`+` next to each other
`/` stacked on top of each other
^ these functions are with the `library(patchwork)` package

