---
title: "Numerical Exploratory Data Analysis"
author: "Lauren Holley"
date: 2026-10-08
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)
library(haven)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

Load the weather data as well.

```{r}
data("weather_df")

weather_df =
  weather_df |>
  mutate(
    month = lubridate::floor_date(date, unit = "month")
  )
```

using these functions to standardize across month categorizations

Now we have everything we need!

```{r}
weather_df |>
  filter(tmax >= 20, tmax <= 30) |>
  ggplot(aes(x = tmax, y= tmin, color = name, shape = name)) +
  geom_point()
```

## `group_by()`

add some groups! 

```{r}
weather_df |>
  group_by(name, month)
```

## `summarize()`

```{r}
weather_df |>
  group_by(name, month) |>
  summarize(count = n())
```


```{r}
weather_df |>
  count(name, month)
```

```{r}
weather_df |>
  group_by(month) |>
  summarize(
    count = n(),
    n_days = n_distinct(date)
    )
```

DONT DO THIS

```{r, eval = FALSE}
weather_df |>
  pull(tmax) |>
  summary()
```

this instead:

```{r}
weather_df |>
  group_by(name) |>
  summarize(
    n = n(),
    mean_tmax = mean(tmax, na.rm = TRUE),
    median_tmax = median(tmax, na.rm = TRUE),
    sd_prcp = sd(prcp, na.rm = TRUE),
    q95_prcp = quantile(prcp, 0.95, na.rm = TRUE)
  ) |>
  knitr::kable(digits = 2)
```

```{r}
weather_df |>
  group_by(name, month) |>
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE)
  ) |>
  pivot_wider(
    names_from = name,
    values_from = mean_tmax
  ) |>
  knitr::kable(digits = 2)
```

This `knitr::kable()` makes the table show up in the markddown -- nice for presenting the data or sharing with others.

What about looking at this

```{r}
weather_df |>
  group_by(name, month) |>
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE)
  ) |>
  ggplot(aes(x = month, y = mean_tmax, color = name)) +
  geom_point() +
  geom_line()
```

```{r}
weather_df |>
  group_by(name) |>
  mutate(center_tmax = tmax - mean(tmax, na.rm = TRUE)) |>
  ggplot(aes(x = date, y = center_tmax, color = name)) +
  geom_point()
```

What about "window" functions...
try to rank things

```{r}
weather_df |>
  group_by(name, month) |>
  mutate(
    temp_rank = min_rank(desc(tmax))
  ) |>
  filter(temp_rank < 2)
```

lead and lag

```{r}
weather_df |>
  group_by(name) |>
  mutate(
    lagged_tmax = lag(tmax),
    temp_change = tmax - lagged_tmax
  ) |>
  summarize(
    mean_temp_change = mean(temp_change, na.rm = TRUE),
    sd_temp_change = sd(temp_change, na.rm = TRUE)
  )
```

```{r}
weather_df |>
  group_by(name) |>
  mutate(
    temp_change = tmax - lag(tmax)
  ) |>
  summarize(
    mean_temp_change = mean(temp_change, na.rm = TRUE),
    sd_temp_change = sd(temp_change, na.rm = TRUE)
  )
```

side note: option shift to select cursor along multiple lines, you can edit multiple lines in the same way

## Revisit some examples

import, clean, tidy, etc pulse data and compute mean and median BDI score at each visit

```{r}
pulse_df =
  read_sas("data/public_pulse_data.sas7bdat") |>
  janitor::clean_names() |>
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    names_prefix = "bdi_score_",
    values_to = "bdi"
  ) |>
  select(id, visit, everything()) |>
  mutate(
    visit = replace(visit, visit == "bl", "00m")
  )

pulse_df |>
  group_by(visit) |>
  summarize(
    mean_bdi = mean(bdi, na.rm = TRUE),
    median_bdi = median(bdi, na.rm = TRUE)
  ) |>
  knitr::kable(digits = 2)
```

in the FAS data, compute mean outcome (ears only) across dose and day of treatment; show in a reader friendly table

```{r}
pups_df = 
  read_csv("data/FAS_pups.csv", skip = 3, na = c("", ".", "NA")) |>
  janitor::clean_names()

litters_df = 
  read_csv("data/FAS_litters.csv", na = c("", ".", "NA")) |>
  janitor::clean_names() |>
  separate(group, into = c("dose", "day_of_tx"), 3)

fas_df =
  left_join(
    pups_df, litters_df, by = "litter_number"
  ) |>
  select(litter_number, dose, day_of_tx, everything()) |>
  drop_na(dose, day_of_tx)

fas_df |>
  group_by(dose, day_of_tx) |>
  summarize(
    mean_ears = mean(pd_ears, na.rm = TRUE)
  ) |>
  pivot_wider(
    names_from = day_of_tx,
    values_from = mean_ears
  )
```

